% vim: foldmethod=marker: 
\documentclass[11pt,twoside,a4paper]{article}

\input{$HOME/math/measureTheory/preamble}

%|--- Author, Title and Date {{{1
\title{Exercise 7 - chapter 4. Book: "Introdução a Medida e Integração"} 
\author{Rafael Polli Carneiro}
\date{August 2021} 
% 1}}}

%|--- Begin Document {{{1
\begin{document}
\maketitle

\begin{definition}
    Let $(\Omega, \mathcal{A}, \mu)$ be a measurable space. Then, we say that 
    any measurable set $A \in \mathcal{A}$ is an atom if
    \begin{equation*}
        \forall B \in \mathcal{A}
        \left( 
            B \subseteq A \implies \mu(B) = 0 \text{ or } \mu(B) = \mu(A).
        \right)
    \end{equation*}
    Thus we say that a masure $\mu$ is non-atomic if it doesn't admit an
    atomic measurable space as element of its $\sigma-$algebra. In other
    words, $\mu$ is non-atomic if
    \begin{equation*}
        \forall A \in \mathcal{A},
        \exists B \in \mathcal{A}\setminus\{\emptyset\}
        \left(
        B \subseteq A \text{ and } 0 < \mu(B) < \mu(A)
        \right)
    \end{equation*}
\end{definition}

Now that we know the meaning of a measure being non-atomic we will proove
the following
\begin{proposition}
    Let $(\Omega, \mathcal{A},\mu)$ be a measurable space and
    $\mu$ a non-atomic measure. Then for any measurable set $A \in \mathcal{A}$    
    such that
    \begin{equation*}
        \mu(A)  < \infty
    \end{equation*}
    we have guaranteed that for any real number $\alpha \in \R$
    satisfying
    \begin{equation*}
        0 < \alpha < \mu(A)        
    \end{equation*}
    there is $B in \mathcal{A}$ where the equality
    \begin{equation*}
        \mu(B)  = \alpha
    \end{equation*}
    holds.
\end{proposition}
\begin{proof}
    The proof for this proposition is quite trick. I will make use of the
    Zorn's Lemma to proceed with the proof 
    (check the section below \ref{sec:zorn}).
\end{proof}

\section {Zorn's Lemma} \label{sec:zorn}
We say that a non empty set $X$ is endowed with a partial order $\leq$    
if
\begin{enumerate}[(i)]
    \item $\leq$ is \textit{reflexive}, which means
    \begin{equation*}
        \forall x \in X ( x \leq x)
    \end{equation*}

    \item $\leq$ is \textit{anti-simmetric}:
    \begin{equation*}
        \forall x,y \in X ( x \leq y, y \leq x \implies x = y )
    \end{equation*}
    \item $\leq$ is \textit{transitive}:
    \begin{equation*}
        \forall x,y,z \in X (x \leq y, y \leq z \implies x \leq z).
    \end{equation*}
\end{enumerate}
Every set $X$ in posse of a partial order is called as a partially ordered   
set. 

Before we adress the Zorn's Lemma we must introduction some notions.
Consider $X$ a partially ordered set and $Y \subseteq X$ a subset.
We say that 
\begin{enumerate}
    \item $x_0 \in X$ is an \textit{upper bound} of $Y$ if
    $\forall y \in Y, y \leq x_0$;

    \item $x_0 \in X$ is a \textit{maximal element} of $X$ if    
    $\forall x \in X, x \leq x_0$;

    \item $Y$ is a \textit{chain} if 
    $\forall x,y \in Y, x \leq y \text{ or } y \leq x$.
\end{enumerate}

Now we are able of stating the Zorn's Lemma, which is a proposition
equivalent to the axiom of choice
\begin{proposition}
    Let $X, \leq$ be a partially ordered set. If any chain
    $Y \subseteq X$ admits an upper bound on $X$, that is
    \begin{equation*}
        \forall Y \subseteq X, \forall y \in Y, \exists y_0 \in X
        \left(
          Y \text{ non empty and it is a chain } \implies y \leq y_0
        \right),
    \end{equation*}
    then $X$ does have a maximal element.  
\end{proposition}

\section{Inducing a partial order into a measurable space}
As usual, consider $(\Omega, \mathcal{A}, \mu)$ a measurable space.
($\Omega$ non empty). We say that two measurable sets $X,Y \in \mathcal{A}$  
are almost equal if 
\begin{equation*}
    \mu( X \setminus Y \cup Y \setminus X ) = 0.
\end{equation*}
From this definition we state the following relation
\begin{equation*}
    \forall X, Y \in \mathcal{A}
    X \sim Y \iff 
    \mu( X \setminus Y \cup Y \setminus X ) = 0.
\end{equation*}
Notice that
\begin{enumerate}[(i)]
   \item $\forall X \in \mathcal{A} \implies X \sim X$;
   
   \item $\forall X, Y \in \mathcal{A}$, clearly $X \sim Y$ 
   and $Y \sim X$;

   \item $\forall X, Y, Z \in \mathcal{A}$ such that
   \begin{equation*}
        X \sim Y \quad \text{and} \quad Y \sim Z
   \end{equation*}
   then 
   \begin{align*}
    \mu( Z \setminus X \cup X \setminus Z )
        & = \mu( [(Z \setminus X) \cup (X \setminus Z)] \cap (Y \cup Y^c) )\\
        & = \mu( [(Z \setminus X)\cap (Y \cup Y^c)]
               \cup 
               [(X \setminus Z) \cap (Y \cup Y^c)] )\\
        & = \mu( [(Z \cap X^c \cap Y) \cup (Z \cap X^c \cap Y^c)]
               \cup 
               [(X \cap Z^c \cap Y)  \cup (X \cap Z^c \cap Y^c)] )\\
        & \leq
            \mu( Z \cap X^c \cap Y)
            + \mu(Z \cap X^c \cap Y^c)
            + \mu(X \cap Z^c \cap Y) 
            + \mu(X \cap Z^c \cap Y^c)\\
        & \leq
            \mu(X^c \cap Y)
            + \mu(Z \cap Y^c)
            + \mu(Z^c \cap Y) 
            + \mu(X \cap Y^c)\\
        & = 0.
   \end{align*}
\end{enumerate}
consequently, we conclude that $\sim$ is an equivalence relation and
therefore we can reduce $\mathcal{A}$ into its equivalence classes  
\begin{equation*}
    \mathcal{A} / \sim = \{ [A]; \; \forall A \in \mathcal{A} \}
\end{equation*}
where
\begin{equation*}
    \forall A \in \mathcal{A}, [A] = \{ B \in \mathcal{A}; \; B \sim A\}.
\end{equation*}

Now we need set operators into the set $\mathcal{A}/\sim$.
Define, for all $[A], [B] \in \mathcal{A}/\sim $:
\begin{equation*}
   [A] \sqcup [B] = [A \cup B],
   \quad
   [A] \sqcap [B] = [A \cap B]
\end{equation*}
and
\begin{equation*}
    [A] \setminus [B] = [A \setminus B]
\end{equation*}

Before we carry on, notice that: 
\begin{itemize}
    \item  $\mathcal{A} / \sim$ is still a $\sigma-$algebra;    
    \item and $\mu$ can be extended to such $\sigma-$algebra.
\end{itemize}

\subsection{ $\mathcal{A} / \sim$ is still a $\sigma-$algebra }

Whenever operating with equivalence relations, one must show
that the operations extended from its base are well definied.

We start by the union.
Let $A_0, A_1, B_0, B_1$ be measurable sets satisfying
\begin{equation*}
    A_0 \sim A_1 \quad \text{ and } \quad B_0 \sim B_1.
\end{equation*}
Let's check what is the value of
\begin{equation*}
    [A_1] \sqcup [B_1] =  [A_1 \cup B_1].
\end{equation*}
We just need to inspect that
\begin{equation*}
    A_1 \cup B_1 \sim A_0 \cup B_0.
\end{equation*}
Let's do it
\begin{align*}
    &\mu\Big( ((A_1 \cup B_1) \setminus ( A_0 \cup B_0) )  \cup  
       ((A_0 \cup B_0) \setminus ( A_1 \cup B_1) ) \Big)\\
	&\qquad=
        \mu( ((A_1 \cup B_1) \cap ( A_0^c \cap B_0^c) ) \cup  
           ((A_0 \cup B_0) \cap ( A_1^c \cap B_1^c) )\\
	&\qquad=
        \mu( ((A_1 \cap A_0^c \cap B_0^c) \cup (B_1 \cap A_0^c \cap B_0^c))
           \cup
           ((A_0 \cap A_1^c \cap B_1^c) \cup (B_0 \cap A_1^c \cap B_1^c))\\
    & \qquad \leq
        \mu( A_1 \cap A_0^c) \mu (B_1 \cap B_0^c)
        +
        \mu (A_0 \cap A_1^c) \mu(B_0 \cap B_1^c))\\
	& \qquad =
        0.
\end{align*}
As a consequence
    \begin{equation*}
    A_1 \cup B_1 \sim A_0 \cup B_0 \implies
    [A_1 \cup B_1] = [ A_0 \cup B_0 ]
    \end{equation*}
and we conclude that the unio is well defined.

For the operator $\sqcap$ we will have the same arguments above and
hence I won't do it. It only rest to check the complementary operation.

To show that taking the complementar is still a well defined operation
we nee to show that
\begin{equation*}
    [A_1] \setminus [B_1] = [A_1 \setminus B_1]
\end{equation*}
is equal to the same above but changing the indexes to $0$.   
Again, we need to show that $A_1 \setminus B_1 \sim A_0 \setminus B_0$.
This is done below
\begin{align*}
    &\mu( (A_1 \setminus B_1) \setminus (A_0 \setminus B_0) 
      \cup
     (A_0 \setminus B_0) \setminus (A_1 \setminus B_1) ) \\
        & \qquad=
        \mu( ((A_1 \cap B_1^c) \cap (A_0 \cap B_0^c)^c)
          \cup
         ((A_0 \cap B_0^c) \cap (A_1 \cap B_1^c)^c) )\\
        & \qquad=
        \mu( ((A_1 \cap B_1^c) \cap (A_0^c \cap B_0))
          \cup
         ((A_0 \cap B_0^c) \cap (A_1^c \cap B_1)) )\\
        & \qquad \leq
         0.
\end{align*}

As a result we have shown that the operations ove classes are well defined.
Now we need to prove that $\mathcal{A}/\sim$  
\end{document}
% 1}}}


